---
title: "THindex"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{THindex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this package, we use the ecology methods to estimate the Tumor Heterogeneity(TH) based on their mutated loci of variant allele frequencis(VAF).

The inferHeterogeneityPlus estimate the TH based on two different methods in the Package [vegan](https://cran.r-project.org/web/packages/vegan/vignettes/diversity-vegan.pdf): 1) **Diveristy indices** and 2) **Taxonomic indices**.

1. Diversity indices

Function diversity finds the most commonly used diversity indices.

$H=-\sum_{i=1}^{S}~p_ilogp_i$      Shannon (1)


$D=\frac{1}{\sum_{i=1}^{S}~p_i^2} \\$       inverse Simpson (2)

Where $p_i$ is the proportion of species $i$ and $S$ is the number of species. For the tumor data, the VAFs of mutated loci in the tumor were assigned to i-th of S bins, and the parameter $p_i$ the proportion of mutated loci belonging to the bins. Here, we set the bin size to 10 (Parameter **region_num** controls the number of bins), yielding enough information to represent the distribution for proprotions of VAFs.

2. Taxonomic diversity.

The simple diveristy above only consider species identity: all species are euqally different. In contrast, taxonimic diveristy indices judge the differences of species. 

$\Delta=\frac{\sum\sum_{i<j}~\omega_ijX_iX_j}{n(n-1)/2}$    Taxonomic diveristy  (3)

$\Delta^*=\frac{\sum\sum_{i<j}~\omega_ijX_iX_j}{\sum\sum_{i<j}~X_iX_j}$      Taxonomic distinctness (4)

These equations give the index values for Taxonomic difference, and summation goes over species $i$ and $j$, and $\omega$ are the taxonomic distances among taxa, $X$ are species abundances, and $n$ is the total abundance for a site.

For the tumor data, the distance of adjacent bins is set 1. For example, if the bins are set 5, then the distance between bin #1 and #5 is 4. If the numbers of occurrences for the 5 bins are (2,4,0,4,2) or (0,2,4,4,3), the former one has higher Taxonomic diversity than the latter one.  



```{r setup, tidy=TRUE}

library(THindex)
library(maftools)

laml.maf <- system.file("extdata", "tcga_laml.maf.gz", package = "maftools")
laml <- read.maf(maf = laml.maf)
TCGA.ab.het <- inferHeterogeneityPlus(maf = laml, vafCol = 'i_TumorVAF_WU', index = "diversity")
print(TCGA.ab.het$diveristy)

TCGA.ab.het1 <- inferHeterogeneityPlus(maf = laml, vafCol = 'i_TumorVAF_WU', index = "taxonomic")
print(TCGA.ab.het1$diveristy)

```
